Notes on modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients

### 2018-06-18 docs/mvfits_paper/mvf.tex
updating doc

\par In practice, \eqn{hlm} is computationally preferable:

@@

%
\section{Methods}
\label{meth}

We begin by ...

% Forward greedy algo
\hspace{1cm}
{\scriptsize
\begin{algorithm}[H]
  %
  \caption{A positive (forward) greedy algorithm, \texttt{PGREEDY}. Note that a required input, $\mathcal{A}$, is a function that takes in a list of basis symbols, and outputs an estimator of fit error. In this setting, $\mathcal{A}$ is assumed to have access to peropheral information, such as the training data.}
  \label{alg:pgreedy}
  %
  \begin{algorithmic}[1]
    %
    \State {\bf Input:} $ \{ \mlam_{bulk} = \text{basis symbols}$, $\mathcal{A} = \text{action}$, $tol=\text{greedy tolerance}\}$
    \vskip 10pt
    \State Define empty list of kept symbols: $\mlam_{kept} = \{\}$
    \State Initialize estimator value and loop boolean: $\epsilon_{last} = \mathrm{inf}$, $done = \text{False}$
    %
    \While{not $done$}
      \State $\epsilon_{min} = \epsilon_{last}$
      %
      \For { $\mlam$ in $\mlam^{bulk}$ }
        \State $\mlam_{trial} = \mlam_{kept} \cup \mlam$ {\hskip0.525in} (add $\mlam$ to $\mlam_{kept}$)
        \State $\epsilon = \mathcal{A}(\mlam_{trial})$ {\hskip0.70in} (action returns fit error)
        \If { $\epsilon < \epsilon_{min}$ }
          \State $\epsilon_{min} = \epsilon$  {\hskip0.825in} (store trial min)
          \State $\mlam_{min} = \mlam_{trial}$
        \EndIf
      \EndFor
      %
      \State $done = |\epsilon_{min}-\epsilon_{last}|<tol$ %{\hskip0.4in} (determine if estimator has changed significantly)
      \If { not $done$ }
        \State $\epsilon_{last} = \epsilon_{min}$
        \State $\mlam_{kept} = \mlam_{kept} \cup \mlam_{min}$ {\hskip0.4in} (update kept symbols)
      \EndIf
      %
    \EndWhile
    \vskip 10pt
    \State {\bf Output:} Greedy Basis, $\{ \mlam_{kept} \}$
    %
  \end{algorithmic}
  %
\end{algorithm}
}

% GMVP
\hspace{1cm}
{\scriptsize
\begin{algorithm}[H]
  %
  \caption{\gmvp{}, a degree tempered stepwise algorithm for multivariate polynomial modeling of scalar data.}
  \label{alg:pgreedy}
  %
  \begin{algorithmic}[1]
    %
    \State {\bf Input:} $ \{ \bf{x}, \bf{y} \}$
    \vskip 10pt
    \State Define action using matrix least squares
    \State Define bulk symbol space using cartesian inner-product
    \State Define list of allowed polynomial degrees
    \For { all degrees }
      \State Select all symbols with degree less than or equal to current
      \State Apply \texttt{PGREEDY} to get symbol subset and estimator val
      \If  { estimator has stalled }
        \State Break
      \EndIf
    \EndFor
    \State Apply negative greedy to get final model
    \vskip 10pt
    \State {\bf Output:} Final model
    %
  \end{algorithmic}
  %
\end{algorithm}
}

>>>>

\par In practice, \eqn{hlm} is computationally efficient:

@@

\section{Methods}
\label{meth}
Within the topic of regression, linear regression has particular advantages:
its matrix based formulation can be computationally efficient, and it does not require initial guesses for model parameters.
Perhaps most intriguingly, the formal series expansions of smooth functions support linear and rational models (e.g. Pade approximants) having application to many datasets.
Here we will develop algorithms for the linear modeling of scalar functions (real or complex valued) of many variables.
\par If we consider a scalar function, $f$, of $N$ variables sampled in $j$, $\vec{x}_j = \{x_{\alpha j}\}_{\alpha=0}^{N}$, then $f(\vec{x})_j$ can be represented (possibly inaccurately) as a sum over $K$ linearly independent basis functions, $\phi_k(\vec{x}_j)$:
  \label{lin1}
  f({\vec{x}})_j \; = \; \sum_{k=0}^{K} \; \mu_{k} \, \phi_{kj}\; .
The central player is \eqn{lin1} is the set of basis coefficients $\mu_{k}$.
Typically, one chooses or derives $\phi_{k}(\vec{x}_j)$ to capture inherent features of $f(\vec{x}_j)$.
With $\phi_{k}(\vec{x}_j)$ assumed to be known, the linear representation (\ceqn{lin1}) lastly defined by the set of $\mu_k$.
\def\vecmu{\vec{\mu}}
\def\vecf{\vec{f}}
\def\hatU{\hat{U}}
\def\hatP{\hat{P}}
\par From here it is useful to note that \eqn{lin1} has a linear homogeneous matrix form. In particular, defining $\hat{U} = \{\phi_{kj}\}$, and $\vec{f} = \{f_j\}$, then $\vec{f} = \hat{U} \; \vec{\mu}$.
This implies that
  \vecmu = \hatP \; \vecf \; ,
where
  \label{pinv1}
  \hatP = \left( \left( \hatU^{-1} \right)^T \hatU^{-1} \right)^{-1}
is the \textit{pseudo-inverse} of $\hatU$.
Equations (\ref{lin1}), and related discussion through \eqn{pinv1} illustrate the most rudimentary solution to the linear modeling problem.
However, there are many ways to expand upon and refine the solution presented thus far.
In the following subsections we will consider two such approaches.
First we will consider the general polynomial modeling of multivariate scalar functions.
This will encompass the \gmvp{} algorithm.
Second, we will build upon the \gmvp{} approach by considering models of rational functions: polynomials divided by polynomilas.
To consider these two approaches in a largely automated way (i.e. where the set of possible basis functions is known, but the select basis functions ultimately used are \textit{learned}), we will make use of the \textit{greedy} algorithm approach.
\subsection{A Generic Greedy Algorithm}
While we most often want a single model for a given dataset (e.g. some approximation of $\vecf$ from numerical calculation or experiment), there are often many more modeling choices than desired.
In particular, if we refer to our set of all possible basis functions as our ``\textit{symbol space}'', then the problem of determining how many, and which basis vectors (i.e. symbols) to use is a problem of combinatoric complexity.
\par A well known method for finding an approximate solution to this problem is the so-called ``greedy'' algorithm.
The premise is this:
We will iteratively construct models with increasing number of symbols.
The process begins by finding the single symbol (basis vector) that yields the most accurate model (in the sense of minimizing least-squares error).
That encompasses the first iteration of a process in which we will greedily add symbols to our model.
In each following iteration, an all remaining symbols are added to the model one at a time, and the single additional symbol which yields the greatest increase in model accuracy is kept for the following iteration.
In this way, a list of optimal model symbols is learned.
The forward process is to end when the model accuracy (or changes thereof) pass a specified threshold.
This rough algorithmic picture is spelled out in more Alg. \ref{alg:pgreedy}.
% \par
% Forward greedy algo
\hspace{1cm}
{\scriptsize
\begin{algorithm}[H]
  \caption{A positive (forward) greedy algorithm, \texttt{PGREEDY}. Note that a required input, $\mathcal{A}$, is a function that takes in a list of basis symbols, and outputs an estimator of fit error. In this setting, $\mathcal{A}$ is assumed to have access to peropheral information, such as the training data.}
  \label{alg:pgreedy}
  \begin{algorithmic}[1]
    \State {\bf Input:} $ \{ \mlam_{bulk} = \text{basis symbols}$, $\mathcal{A} = \text{action}$, $tol=\text{greedy tolerance}\}$
    \vskip 10pt
    \State Define empty list of kept symbols: $\mlam_{kept} = \{\}$
    \State Initialize estimator value and loop boolean: $\epsilon_{last} = \mathrm{inf}$, $done = \text{False}$
    \While{not $done$}
      \State $\epsilon_{min} = \epsilon_{last}$
      \For { $\mlam$ in $\mlam^{bulk}$ }
        \State $\mlam_{trial} = \mlam_{kept} \cup \mlam$ {\hskip0.525in} (add $\mlam$ to $\mlam_{kept}$)
        \State $\epsilon = \mathcal{A}(\mlam_{trial})$ {\hskip0.70in} (action returns fit error)
        \If { $\epsilon < \epsilon_{min}$ }
          \State $\epsilon_{min} = \epsilon$  {\hskip0.825in} (store trial min)
          \State $\mlam_{min} = \mlam_{trial}$
        \EndIf
      \EndFor
      \State $done = |\epsilon_{min}-\epsilon_{last}|<tol$ %{\hskip0.4in} (determine if estimator has changed significantly)
      \If { not $done$ }
        \State $\epsilon_{last} = \epsilon_{min}$
        \State $\mlam_{kept} = \mlam_{kept} \cup \mlam_{min}$ {\hskip0.4in} (update kept symbols)
      \EndIf
    \EndWhile
    \vskip 10pt
    \State {\bf Output:} Greedy Basis, $\{ \mlam_{kept} \}$
  \end{algorithmic}
\end{algorithm}
}

%
\subsection{Greedy Multivariate Polynomial Fitting}
%
The preceeding ...
% GMVP
\hspace{1cm}
{\scriptsize
\begin{algorithm}[H]
  %
  \caption{\gmvp{}, a degree tempered stepwise algorithm for multivariate polynomial modeling of scalar data.}
  \label{alg:pgreedy}
  %
  \begin{algorithmic}[1]
    %
    \State {\bf Input:} $ \{ \bf{x}, \bf{y} \}$
    \vskip 10pt
    \State \grey{\# Define action using matrix least squares}
    \State $\mathcal{A}$: $\epsilon = \hat{U}$
    \State Define bulk symbol space using cartesian inner-product
    \State Define list of allowed polynomial degrees
    \For { all degrees }
      \State Select all symbols with degree less than or equal to current
      \State Apply \texttt{PGREEDY} to get symbol subset and estimator val
      \If  { estimator has stalled }
        \State Break
      \EndIf
    \EndFor
    \State Apply negative greedy to get final model
    \vskip 10pt
    \State {\bf Output:} Final model
    %
  \end{algorithmic}
  %
\end{algorithm}
}
